---
title: "Mana Simulation"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    source_code: embed
runtime: shiny
---

```{r global, include=FALSE}
library(shiny); library(stringr); library(data.table);

lands_to_deck <- function(x, size=60) {
  x <- x %>%  str_replace_all("\n", " ") %>% # replace newlines with spaces
    str_replace_all("\\s+", " ") %>% # shorten any long whitespaces
    trimws() %>% 
    str_split(" ") # divide the string into elements
  x <- x[[1]] # convert from list to vector
  # x == c("10", "R", "10", "G", "4", "RG")
  
  mana <- c()
  for (i in seq(1, length(x), by=2)) {
    mana <- c(mana, rep(x[i+1], x[i])) # convert the counts to repetitions
  }
  
  deck <- c(mana, rep("x", size - length(mana))) # fill in deck with spelkls until 60 cards
  deck
}

rlanddrop <- function(deck, m=1000, colors=F, mul=T, draw_cards=0) {
  size <- length(deck)
  mana_df <- data.frame(t1=numeric(), t2=numeric(), t3=numeric(), t4=numeric(), t5=numeric())
  
  ### For each color, check if it's in the deck. 
  ### If not, then has_color will be F and then the function will ignore making a df for that color.
  if (colors==T){
    has_white <- F; has_blue <- F; has_black <- F; has_red <- F; has_green <- F;
    
    if (sum(str_detect(deck, "W")) > 0){
      white_df <- data.frame(t1=numeric(), t2=numeric(), t3=numeric(), t4=numeric(), t5=numeric())
      has_white <- T
    }
    if (sum(str_detect(deck, "U")) > 0){
      blue_df <- data.frame(t1=numeric(), t2=numeric(), t3=numeric(), t4=numeric(), t5=numeric())
      has_blue <- T
    }
    if (sum(str_detect(deck, "B")) > 0){
      black_df <- data.frame(t1=numeric(), t2=numeric(), t3=numeric(), t4=numeric(), t5=numeric())
      has_black <- T
    }
    if (sum(str_detect(deck, "R")) > 0){
      red_df <- data.frame(t1=numeric(), t2=numeric(), t3=numeric(), t4=numeric(), t5=numeric())
      has_red <- T
    }
    if (sum(str_detect(deck, "G")) > 0){
      green_df <- data.frame(t1=numeric(), t2=numeric(), t3=numeric(), t4=numeric(), t5=numeric())
      has_green <- T
    }
  }
  
  ### The simulation begins.
  for (r in 1:m){ 
    # Get list of hands by turn
    deck <- deck[sample(1:size)]
    hands <- list(deck[1:7])
    
    ### Mulligan function
    if (mul==T){
      num_lands <- sum(!str_detect(hands[[1]], "x"))
      
      if (num_lands < 2 | num_lands > 5) {
        deck <- deck[sample(1:size)]
        hands <- list(deck[1:6])
        num_lands <- sum(!str_detect(hands[[1]], "x"))
        
        if (num_lands < 2 | num_lands > 4){
          deck <- deck[sample(1:size)]
          hands <- list(deck[1:5])
          num_lands <- sum(!str_detect(hands[[1]], "x"))
          
          if (num_lands == 0 | num_lands == 5){
            deck <- deck[sample(1:size)]
            hands <- list(deck[1:4])
          }
        }
        
        # Because mulligan once or more, scry a land to top and spell to bottom.
        scry_card <- deck[length(hands[[1]])+1]
        if (scry_card == "x") {
          deck[size+1] <- scry_card
          deck <- deck[-(length(hands[[1]])+1)]
        }
      }
    }
    ### End Mulligan
    
    ### Drawing one or more cards on first turn
    if (draw_cards > 0){
      hands <- list(deck[1:(length(hands[[1]]) + draw_cards)])
    }
    ### End draw additional cards
    
    ### Get hand for first 5 turns, drawing 1 card per turn.
    starting_hand_size <- length(hands[[1]])
    
    for (i in 1:4){
      hands <- c(hands, list(deck[1:(starting_hand_size+i)]))
    }
    ### Use hands to get probabilities for colorless land drops
    # Calculate whether a land drop was missed by that turn for each turn.
    # mana_df will be aggregated and returned as output.
    mana_row <- c()
    
    for (i in 1:5){
      mana_row <- cbind(mana_row, sum(!str_detect(hands[[i]], "x")) >= i)
    }
    
    mana_df[r,] <- mana_row
    
    ### Use hands to get probabilities for colorful land drops
    if (colors==T){
      if (has_white==T){
        color_row <- c()
        for (i in 1:5){
          color_row <- cbind(color_row, sum(str_detect(hands[[i]], "W")) >= i)
        }
        white_df[r,] <- color_row
      }
      if (has_blue==T){
        color_row <- c()
        for (i in 1:5){
          color_row <- cbind(color_row, sum(str_detect(hands[[i]], "U")) >= i)
        }
        blue_df[r,] <- color_row
      }
      if (has_black==T){
        color_row <- c()
        for (i in 1:5){
          color_row <- cbind(color_row, sum(str_detect(hands[[i]], "B")) >= i)
        }
        black_df[r,] <- color_row
      }
      if (has_red==T){
        color_row <- c()
        for (i in 1:5){
          color_row <- cbind(color_row, sum(str_detect(hands[[i]], "R")) >= i)
        }
        red_df[r,] <- color_row
      }
      if (has_green==T){
        color_row <- c()
        for (i in 1:5){
          color_row <- cbind(color_row, sum(str_detect(hands[[i]], "G")) >= i)
        }
        green_df[r,] <- color_row
      }
    }
  }
  
  ### Aggregate finished hands for colorless mana
  mana_df <- data.table(mana_df)
  mana_df <- mana_df[, j = list(t1=mean(t1), t2=mean(t2), t3=mean(t3), t4=mean(t4), t5=mean(t5))]
  returned_list <- list(any_color = mana_df)
  
  ### Aggregate finished hands for colorful mana
  if (colors==T){
    if (has_white==T) {
      white_df <- data.table(white_df)
      white_df <- white_df[, j = list(t1=mean(t1), t2=mean(t2), t3=mean(t3), t4=mean(t4), t5=mean(t5))]
      returned_list <- c(returned_list, list(white = white_df))
    }
    if (has_blue==T) {
      blue_df <- data.table(blue_df)
      blue_df <- blue_df[, j = list(t1=mean(t1), t2=mean(t2), t3=mean(t3), t4=mean(t4), t5=mean(t5))]
      returned_list <- c(returned_list, list(blue = blue_df))
    }
    if (has_black==T) {
      black_df <- data.table(black_df)
      black_df <- black_df[, j = list(t1=mean(t1), t2=mean(t2), t3=mean(t3), t4=mean(t4), t5=mean(t5))]
      returned_list <- c(returned_list, list(black = black_df))
    }
    if (has_red==T) {
      red_df <- data.table(red_df)
      red_df <- red_df[, j = list(t1=mean(t1), t2=mean(t2), t3=mean(t3), t4=mean(t4), t5=mean(t5))]
      returned_list <- c(returned_list, list(red = red_df))
    }
    if (has_green==T) {
      green_df <- data.table(green_df)
      green_df <- green_df[, j = list(t1=mean(t1), t2=mean(t2), t3=mean(t3), t4=mean(t4), t5=mean(t5))]
      returned_list <- c(returned_list, list(green = green_df))
    }
  }
  returned_list
}

prob_table <- function(landsindeck, size, extracards, numtrials){
  deck <- lands_to_deck(landsindeck, size)
  x <- rlanddrop(deck, colors=T, draw_cards=extracards, m=numtrials)

  y <- data.frame()
  for (i in 1:length(x)){
    y <- rbind(y,x[[i]])
  }
  
  y <- cbind(color=names(x),y)
  colnames(y) <- c("Color", "turn 1",
                   "turn 2", "turn 3",
                   "turn 4", "turn 5")
  
  y
}
```

Sidebar {.sidebar}
===

```{r}
textAreaInput("lands_in_deck", "Number and type of lands in your deck", value="10 R\n10 G\n4 RG", height='200px')

numericInput("num_cards", "Number of cards in deck", value=60, min=0)

numericInput("extra_cards", "Extra cards drawn on turn 1", value=0, min=0)

selectInput("num_trials", "Simulation Accuracy (nbr of trials)", choices=c("Medium (1000)"=1000, "High (2000)"=2000, "Almost Exact (5000)"=5000), selected=2000)

actionButton("run_simulation", "Run Simulation")
```



<div style='padding: 3px; width: 220px; word-wrap: break-word;'>

</div>

Page 1
=====

Column {data-width=600}
-----------------------------------------
### Instructions
This app calculates (through simulation) the probabilities of not missing any land drops by turns 1 through 5. 

To use the app, input the counts and types of lands in your deck in the box on the left. Number, space, color, newline. W U B R G C stand for white, blue, black, red, green and colorless. Lands that produce multiple colors are also allowed. Then press "Run Simulation."


Optional: Drawing instead of playing first can be simulated by setting "Extra Cards" to 1. The simulation accuracy can also be improved at the cost of runtime.

Link to paper about this simulation:
<a href=https://darrensstuffdrawer.blogspot.com/2018/06/mtg-optimal-land-counts-for-consistent.html target="_blank">https://darrensstuffdrawer.blogspot.com/2018/06/mtg-optimal-land-counts-for-consistent.html</a>


Link to Channel Fireball article that this project is based on:
<a href=https://channelfireball.com/articles/how-many-lands-do-you-need-to-consistently-hit-your-land-drops/ target="_blank">https://channelfireball.com/articles/how-many-lands-do-you-need-to-consistently-hit-your-land-drops/</a>

### Probability of not missing a land drop by turns 1 through 5:
```{r}
renderTable({
  input$run_simulation
  isolate({
      prob_table(input$lands_in_deck, input$num_cards, input$extra_cards, input$num_trials)
  })
})
```


